---
output:
  word_document: default
  html_document: default
---
```{r include = FALSE}
library(tidyverse)
library(tidymodels)
library(e1071)
library(ROCR)
library(caret)
library(skimr)
library(GGally)
library(gridExtra)
library(vip)
library(corrplot)
library(mice) #package for imputation
library(VIM) #visualizing missingness
library(ranger) #for random forests
library(randomForest)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(modelr)
```

```{r}
heart <- read_csv("heart.csv")
```

```{r}
str(heart)
summary(heart)
```

```{r}
skim(heart)
```


```{r}
heart <- heart %>%
  mutate(Sex = as_factor(Sex)) %>%
  mutate(ChestPainType = as_factor(ChestPainType)) %>%
  mutate(RestingECG = as_factor(RestingECG)) %>%
  mutate(ExerciseAngina = as_factor(ExerciseAngina)) %>%
  mutate(ST_Slope = as_factor(ST_Slope)) %>%
  mutate(FastingBS = as.factor(FastingBS)) %>%
  mutate(HeartDisease = as_factor(HeartDisease)) %>%
  mutate(HeartDisease = fct_recode(HeartDisease, "No" = "0", "Yes" = "1")) %>%
  filter(Cholesterol > 0)
str(heart)
```

```{r}
ggplot(data = heart, aes(x = Age)) +
  geom_histogram(color = "darkblue", fill = "lightblue") +
  labs(title = "Age Histogram Plot", x = "Age", y = "Count") +
  theme(plot.title = element_text(hjust=0.5)) + 
  theme_minimal()
```

```{r}
ggplot(data = heart, aes(x = Sex)) +
  stat_count(fill = "steelblue") +
  labs(title = "Sex Bar Plot", x = "Sex", y = "Count") +
  theme(plot.title = element_text(hjust=0.5)) +
  theme_minimal()
```

```{r}
heart$ChestPainType %>% summary()

ggplot(data = heart, aes(x = ChestPainType)) +
  stat_count(fill = "steelblue") +
  labs(title = "Chest Pain Type Bar Plot", x = "Chest Pain Type", y = "Count") +
  theme_minimal()
```

```{r}
heart$RestingBP %>% summary()

ggplot(data = heart, aes(x = RestingBP)) +
  geom_histogram(color = "darkblue", fill = "lightblue") +
  labs(title = "Resting Blood Pressure Histogram Plot", x = "Resting Blood Pressure", y = "Count") +
  theme_minimal()
```

```{r}
heart$Cholesterol %>% summary()

ggplot(data = heart, aes(x = Cholesterol)) +
  geom_histogram(color = "darkblue", fill = "lightblue") +
  labs(title = "Serum Cholesterol Histogram Plot", x = "Serum Cholesterol", y = "Count") +
  theme_minimal()
```
```{r}
heart$FastingBS %>% summary()

ggplot(data = heart, aes(x = FastingBS)) +
  stat_count(fill = "steelblue") +
  labs(title = "Fasting Blood Sugar Bar Plot", x = "Fasting Blood Sugar", y = "Count") +
  theme_minimal()
```
```{r}
heart$RestingECG %>% summary()

ggplot(data = heart, aes(x = RestingECG)) +
  stat_count(fill = "steelblue") +
  labs(title = "Resting Electrocardiogram Results Bar Plot", x = "Resting Electrocardiogram Results", y = "Count") +
  theme_minimal()
```

```{r}
heart$MaxHR %>% summary()

ggplot(data = heart, aes(x = MaxHR)) +
  geom_histogram(color = "darkblue", fill = "lightblue") +
  labs(title = "Maximum Heart Rate Histogram Plot", x = "Maximum Heart Rate", y = "Count") +
  theme_minimal()
```
```{r}
heart$ExerciseAngina %>% summary()

ggplot(data = heart, aes(x = ExerciseAngina)) +
  stat_count(fill = "steelblue") +
  labs(title = "Exercise - Induced Angina Bar Plot", x = "Exercise - Induced Angina Bar Plot", y = "Count") +
  theme_minimal()
```
```{r}
heart$Oldpeak %>% summary()

ggplot(data = heart, aes(x = Oldpeak)) +
  geom_histogram(color = "darkblue", fill = "lightblue") +
  labs(title = "Numeric Value Measured in Depression Histogram Plot", x = "Numeric Value Measured in Depression", y = "Count") +
  theme_minimal()
```
```{r}
heart$ST_Slope %>% summary()

ggplot(data = heart, aes(x = ST_Slope)) +
  stat_count(fill = "steelblue") +
  labs(title = "ST Slope Bar Plot", x = "ST Slope Bar Plot", y = "Count") +
  theme_minimal()
```
```{r}
heart$HeartDisease %>% summary()

ggplot(data = heart, aes(x = HeartDisease)) +
  stat_count(fill = "steelblue") +
  labs(title = "Heart Disease Bar Plot", x = "Heart Disease Bar Plot", y = "Count") +
  theme_minimal()
```


```{r}
p1 = ggplot(heart, aes(x = Sex, fill = HeartDisease)) + geom_bar(position = "fill")
p2 = ggplot(heart, aes(x = RestingECG, fill = HeartDisease)) + geom_bar(position = "fill")
p3 = ggplot(heart, aes(x = ExerciseAngina, fill = HeartDisease)) + geom_bar(position = "fill")
p4 = ggplot(heart, aes(x = ST_Slope, fill = HeartDisease)) + geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4, ncol = 2)
```

```{r}
p1 = ggplot(heart, aes(x = HeartDisease, y = Age)) + geom_boxplot()
p2 = ggplot(heart, aes(x = HeartDisease, y = RestingBP)) + geom_boxplot()
p3 = ggplot(heart, aes(x = HeartDisease, y = Cholesterol)) + geom_boxplot()
p4 = ggplot(heart, aes(x = HeartDisease, y = Oldpeak)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4, ncol = 2)
```

```{r}
p1 = ggplot(heart, aes(x = FastingBS, fill = HeartDisease)) + geom_bar(position = "fill")
p2 = ggplot(heart, aes(x = ExerciseAngina, fill = HeartDisease)) + geom_bar(position = "fill")
grid.arrange(p1,p2, ncol = 2)
```

```{r}
set.seed(123) 
heart_split = initial_split(heart, prop = 0.75, strata = HeartDisease) #75% in training
train = training(heart_split)
test = testing(heart_split)
```

###Logistic Regression 
```{r}
heart_model = 
  logistic_reg(mode = "classification") %>% 
  set_engine("glm") 

heart_recipe = recipe(HeartDisease ~., train)

logreg_wf = workflow() %>%
  add_recipe(heart_recipe) %>% 
  add_model(heart_model)

heart_fit = fit(logreg_wf, train)
```

### Heart Model Fit
```{r}
summary(heart_fit$fit$fit$fit)
```

### Predictions
```{r}
predictions = predict(heart_fit, train, type="prob") #develop predicted probabilities
head(predictions)
```

### Prediction for Yes
```{r}
predictions = predict(heart_fit, train, type="prob")[2]
head(predictions)
```

### ROC Curve
```{r}
ROCRpred = prediction(predictions, train$HeartDisease)

ROCRperf = performance(ROCRpred, "tpr", "fpr")
plot(ROCRperf, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1), text.adj=c(-0.2,1.7))
```

### AUC Value
```{r}
as.numeric(performance(ROCRpred, "auc")@y.values)
```
### Sensitivity, Specificity & Cutoff
```{r}
opt.cut = function(perf, pred){
    cut.ind = mapply(FUN=function(x, y, p){
        d = (x - 0)^2 + (y-1)^2
        ind = which(d == min(d))
        c(sensitivity = y[[ind]], specificity = 1-x[[ind]], 
            cutoff = p[[ind]])
    }, perf@x.values, perf@y.values, pred@cutoffs)
}
print(opt.cut(ROCRperf, ROCRpred))
```
### Use cutoff for best value in the confusion matrix
```{r}
#confusion matrix
#The "No" and "Yes" represent the actual values
#The "FALSE" and "TRUE" represent our predicted values
t1 = table(train$HeartDisease,predictions > .4539403)
t1
```
### Accuracy
```{r}
((t1[1,1]+t1[2,2])/nrow(train)) * 100
```
### Apply to test data
```{r}
testpredlog = predict(heart_fit,test)
head(testpredlog)
```
### Here we wave an accuracy of 88.2%
### 94 patients were correctly classified as not having heart disease
### 18 patients were wrongly classified as having heart disease that did not have heart disease
### 4 Patients were wrongly classified as not having heart disease that had heart disease
### 71 patients were correctly classified as having heart disease
```{r}
confusionMatrix(testpredlog$.pred_class, as_factor(test$HeartDisease),
                positive = "Yes")
```


```{r}
heart_model2 = 
  logistic_reg(mode = "classification") %>% #note the use of logistic_reg and mode = "classification"
  set_engine("glm") #standard logistic regression engine is glm

heart_recipe2 = recipe(HeartDisease ~ Sex + ChestPainType + Cholesterol + FastingBS + ExerciseAngina + Oldpeak + ST_Slope, train)

logreg_wf2 = workflow() %>%
  add_recipe(heart_recipe2) %>% 
  add_model(heart_model2)

heart_fit2 = fit(logreg_wf2, train)
```

```{r}
summary(heart_fit2$fit$fit$fit)
```

```{r}
predictions = predict(heart_fit2, train, type="prob") #develop predicted probabilities
head(predictions)
```

```{r}
predictions = predict(heart_fit2, train, type="prob")[2]
head(predictions)
```

```{r}
ROCRpred = prediction(predictions, train$HeartDisease)

ROCRperf = performance(ROCRpred, "tpr", "fpr")
plot(ROCRperf, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1), text.adj=c(-0.2,1.7))
```

```{r}
as.numeric(performance(ROCRpred, "auc")@y.values)
```

```{r}
opt.cut = function(perf, pred){
    cut.ind = mapply(FUN=function(x, y, p){
        d = (x - 0)^2 + (y-1)^2
        ind = which(d == min(d))
        c(sensitivity = y[[ind]], specificity = 1-x[[ind]], 
            cutoff = p[[ind]])
    }, perf@x.values, perf@y.values, pred@cutoffs)
}
print(opt.cut(ROCRperf, ROCRpred))
```

```{r}
#confusion matrix
#The "No" and "Yes" represent the actual values
#The "FALSE" and "TRUE" represent our predicted values
t1 = table(train$HeartDisease,predictions > .5491748)
t1
```

```{r}
((t1[1,1]+t1[2,2])/nrow(train)) * 100
```

```{r}
testpredlog = predict(heart_fit2,test)
head(testpredlog)
```

```{r}
confusionMatrix(testpredlog$.pred_class, as_factor(test$HeartDisease),
                positive = "Yes")
```
```{r}
heartpredlog = predict(heart_fit2,heart)
head(heartpredlog)

predictions = predict(heart_fit2, heart, type="prob")[2]
head(predictions)
```
### Add predictions to a new data set
```{r}
new_heart = cbind(heart, predictions, heartpredlog)
```

### Random Forests
```{r}
heart_recipe3 = recipe(HeartDisease ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

heart_model3 = rand_forest() %>% 
  set_engine("ranger", importance = "permutation") %>%
  set_mode("classification")

heart_wflow3 = 
  workflow() %>% 
  add_model(heart_model3) %>% 
  add_recipe(heart_recipe3)

set.seed(123)
heart_fit3 = fit(heart_wflow3, train)
```

```{r}
trainpredrf = predict(heart_fit3, train)
head(trainpredrf)
```

```{r}
confusionMatrix(trainpredrf$.pred_class, as_factor(train$HeartDisease), 
                positive = "Yes")
```

### Variable Importance, ST_Slope_Flat explains 8% of the variance
```{r}
heart_fit3 %>% pull_workflow_fit() %>% vip(geom = "point")
```

```{r}
testpredrf = predict(heart_fit3, test)
head(testpredrf)
```

```{r}
confusionMatrix(testpredrf$.pred_class, as_factor(test$HeartDisease), 
                positive = "Yes")
```
### Decision Tree
```{r}
heart_recipe4 = recipe(HeartDisease ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

heart_model4 = decision_tree() %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")

heart_wflow4 = 
  workflow() %>% 
  add_model(heart_model4) %>% 
  add_recipe(heart_recipe4)

heart_fit4 = fit(heart_wflow4, train)
heart_fit4 %>%
  extract_fit_parsnip() %>%
  pluck("fit")  
```

```{r}
tree = heart_fit4 %>% 
  pull_workflow_fit() %>% 
  pluck("fit")
```

```{r}
fancyRpartPlot(tree, tweak=1.5)
```

```{r}
heart_fit4$fit$fit$fit$cptable
```

